---
title: 'Peer-graded Assignment: Course Project 1'
author: "saurabh"
date: "June 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### This document is created as part of the assigment requirements of course reproducible-research- Week 2 Peer-graded Assignment: Course Project 1 .

The [dataset](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) being used for this assigment is of size 56 kb, and has below listed variables

* steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
* date: The date on which the measurement was taken in YYYY-MM-DD format
* interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

Load the required libraries
```{r load_libraries}
library(ggplot2)
library("lubridate")
library(scales)
```

Lets download and unzip the dataset and store into activityData

```{r download_data}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile="Factivity.zip")
unzip("Factivity.zip")
activityData <- read.csv("activity.csv", sep = ",")
```

Now, we should get an overview of the available data

```{r data_summary}
class(activityData)
dim(activityData)
str(activityData)
head(activityData)
tail(activityData)
```


Lets check the data type for date field
```{r date_datatype}
class(activityData$date)
```

We will now convert factor date field to data class date

```{r convertDate}
activityData$date <- as.Date(activityData$date, format = "%Y-%m-%d")
class(activityData$date)
```

We can now remove the NA values,

```{r rm_NAvalues}
activityData <- na.omit(activityData)
```

Now to plot Histogram of the total number of steps taken each day, we will aggregate steps by date, and store the aggreated data in aggdata

```{r aggdata}
aggdata <- setNames(aggregate(activityData[,1], by=list(activityData$date), FUN=sum, na.rm=TRUE), c("Date", "Sum_steps"))
head(aggdata)
tail(aggdata)
hist(aggdata$Sum_steps, main = "Histogram of the total number of steps taken each day", xlab = "total number of steps taken each day")
```

Mean and Median of total number of steps taken in each day are:

```{r mean_median}
mean(aggdata$Sum_steps)
median(aggdata$Sum_steps)
```

Time series plot of the average number of steps taken

```{r}
meandata <- setNames(aggregate(activityData[,1], by=list(activityData$date), FUN=mean, na.rm=TRUE), c("date", "mean_steps"))
ggplot(meandata, aes(date, mean_steps)) + geom_line() + xlab("") + ylab("Mean daily steps") + ggtitle("Time series plot of the average number of steps taken")
```

Now to calculate "The 5-minute interval that, on average, contains the maximum number of steps" we fill aggregate mean of activityData by 5 minute interval

```{r}
meandataall<- setNames(aggregate(activityData[,1], by=list(activityData$interval), FUN=mean, na.rm=TRUE), c("interval","mean_steps"))
meandataall[which.max(meandataall[,2]),]
```

From the above output, we can see that interval 835 is having the maximum number os steps.

To answer rest of questions, lets get the initial dataframe with NA values,
```{r}
activityData <- read.csv("activity.csv", sep = ",")
activityData$date <- as.Date(activityData$date, format = "%Y-%m-%d")
```

Now, we will use Hmisc package and impute to impute missing data with mean value,

```{r}
library("Hmisc")
head(activityData)
activityData$steps <- with(activityData, impute(steps, mean))
head(activityData)
```

Now, lets plot Histogram of the total number of steps taken each day after missing values are imputed
```{r }
aggdata <- setNames(aggregate(activityData[,1], by=list(activityData$date), FUN=sum, na.rm=TRUE), c("Date", "Sum_steps"))
head(aggdata)
tail(aggdata)
hist(aggdata$Sum_steps, main = "Histogram of the total number of steps taken each day", xlab = "total number of steps taken each day")
```

Finally, we will Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
meandata <- setNames(aggregate(activityData[,1], by=list(activityData$date), FUN=mean, na.rm=TRUE), c("date", "mean_steps"))
meandata$day <- wday(meandata$date, label = TRUE)
meandata$week <- ifelse(weekdays(meandata$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
meandata$week <- factor(meandata$week,labels=c("weekday","weekend"))
class(meandata$week)
ggplot(meandata, aes(x=date, y=mean_steps, colour=week)) + 
  geom_path() + 
  geom_point() + scale_x_date(breaks=date_breaks("1 month")) +
  facet_grid(week ~.) +
  theme(legend.position="none")
```
